---
title: "📊 Data Pre-Processing Platform Analysis Report"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: 
      collapsed: false
      smooth_scroll: true
    number_sections: true
    theme: united
    code_folding: hide
    df_print: paged
    self_contained: true
    css: |
      .main-container {
        max-width: 1200px;
        margin-left: auto;
        margin-right: auto;
      }
      h1, h2, h3, h4 {
        margin-top: 24px;
        margin-bottom: 16px;
        font-weight: 600;
        line-height: 1.25;
      }
      ul, ol {
        padding-left: 2em;
        margin-bottom: 16px;
      }
      li {
        margin-bottom: 8px;
        display: list-item;
      }
      li ul, li ol {
        margin-top: 8px;
        margin-left: 1em;
      }
      ul li {
        list-style-type: disc;
      }
      ul li li {
        list-style-type: circle;
      }
      ul li li li {
        list-style-type: square;
      }
      ol li {
        list-style-type: decimal;
      }
      ol li li {
        list-style-type: lower-alpha;
      }
      ol li li li {
        list-style-type: lower-roman;
      }
      p {
        margin-bottom: 16px;
      }
      pre {
        margin-bottom: 16px;
      }
      .table-container {
        overflow-x: auto;
        margin-bottom: 20px;
      }

params:
  processed_data: NULL
  original_data: NULL
  filename: NULL
  preprocessing_history: NULL
  numeric_vars: NULL
  categorical_vars: NULL
  script_dir: NULL
  include_plots: TRUE
  include_stats: TRUE
  include_preprocessing: TRUE
  include_correlation: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")

# Load required libraries
suppressMessages({
  library(ggplot2)
  library(dplyr)
  library(knitr)
  library(kableExtra)
  library(jsonlite)
})

# Webapp color palette (matching VisualizationDisplay.tsx)
webapp_colors <- c(
  "#2563eb", "#16a34a", "#f97316", "#0ea5e9", "#ef4444",
  "#8b5cf6", "#14b8a6", "#f59e0b", "#06b6d4", "#ec4899"
)
```

# Data Pre-Processing and Visualization Report

This report provides a detailed analysis of the data preprocessing steps and visualizations for the **Data Pre-Processing Platform** tool.

## Dataset Overview

### Dataset Information

```{r results="asis"}
# Display dataset name
if (!is.null(params$filename)) {
  cat("<p><strong>Dataset Name:</strong>", params$filename, "</p>")
} else {
  cat("<p><strong>Dataset Name:</strong> Not specified</p>")
}

# Add analysis date
cat("<p><strong>Analysis Date:</strong>", format(Sys.Date(), "%B %d, %Y"), "</p>")
```

### Data Preview

```{r}
# Display first 10 rows of processed data
if (!is.null(params$processed_data) && nrow(params$processed_data) > 0) {
  kable(head(params$processed_data, 10),
    caption = "First 10 rows of dataset"
  ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = TRUE
    ) %>%
    scroll_box(width = "100%", height = "300px")
} else {
  cat("<p>No data available to display.</p>")
}
```

### Dataset Structure

```{r results="asis"}
# Display data dimensions
if (!is.null(params$processed_data) && ncol(params$processed_data) > 0) {
  cat(
    "<p><strong>Dataset Dimensions:</strong>",
    format(nrow(params$processed_data), big.mark = ","),
    "rows ×", ncol(params$processed_data), "columns</p>\n\n"
  )
}
```

## Data Preprocessing

### Preprocessing Operations

```{r}
# Display preprocessing operations from history
if (params$include_preprocessing && !is.null(params$preprocessing_history)) {
  # Handle case where preprocessing_history might be a list or data frame
  history_list <- if (is.data.frame(params$preprocessing_history)) {
    # Convert data frame rows to list if needed
    lapply(1:nrow(params$preprocessing_history), function(i) as.list(params$preprocessing_history[i, ]))
  } else {
    params$preprocessing_history
  }

  if (length(history_list) > 0) {
    # Extract preprocessing operations
    prep_ops <- data.frame(
      Operation = character(),
      Input = character(),
      Result = character(),
      Method = character(),
      stringsAsFactors = FALSE
    )

    for (i in seq_along(history_list)) {
      op <- history_list[[i]]

      # Access list elements safely
      action <- if (!is.null(op$action)) op$action else if (!is.null(op[["action"]])) op[["action"]] else NA

      if (!is.na(action) && action %in% c("Imputation", "Encoding", "Binning", "Data Reduction", "Text Transformation")) {
        inputs <- if (!is.null(op$inputs)) paste(op$inputs, collapse = ", ") else "-"
        outputs <- if (!is.null(op$outputs)) paste(op$outputs, collapse = ", ") else "-"
        method <- if (!is.null(op$method)) op$method else "-"

        prep_ops <- rbind(prep_ops, data.frame(
          Operation = action,
          Input = inputs,
          Result = outputs,
          Method = method,
          stringsAsFactors = FALSE
        ))
      }
    }

    if (nrow(prep_ops) > 0) {
      kable(prep_ops, caption = "Applied Preprocessing Operations") %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
    } else {
      cat("No preprocessing operations were recorded.")
    }
  } else {
    cat("No preprocessing operations were recorded.")
  }
} else {
  cat("No preprocessing operations were recorded.")
}
```

### Missing Values Analysis

```{r}
# Check for missing values
if (params$include_stats && !is.null(params$processed_data)) {
  missing_counts <- sapply(params$processed_data, function(x) sum(is.na(x)))
  missing_counts <- missing_counts[missing_counts > 0]

  if (length(missing_counts) > 0) {
    cat("#### Columns with Missing Values\n\n")
    for (col_name in names(missing_counts)) {
      missing_count <- missing_counts[col_name]
      missing_pct <- round(100 * missing_count / nrow(params$processed_data), 2)
      cat("*", col_name, ":", missing_count, "values (", missing_pct, "%)\n\n", sep = " ")
    }
  } else {
    cat("**No missing values** found in the dataset.\n\n")
  }
}
```

## Data Exploration

### Summary Statistics

```{r}
if (params$include_stats && !is.null(params$processed_data) && ncol(params$processed_data) > 0) {
  # Create summary statistics table
  summary_df <- data.frame(
    Column = names(params$processed_data),
    Class = sapply(params$processed_data, function(x) {
      if (is.numeric(x)) "Numerical" else "Categorical"
    }),
    Missing = sapply(params$processed_data, function(x) sum(is.na(x))),
    Missing_Percent = sapply(params$processed_data, function(x) {
      sprintf("%.1f%%", (sum(is.na(x)) / length(x)) * 100)
    }),
    Min = sapply(params$processed_data, function(x) {
      if (is.numeric(x)) round(min(x, na.rm = TRUE), 2) else "-"
    }),
    Median = sapply(params$processed_data, function(x) {
      if (is.numeric(x)) round(median(x, na.rm = TRUE), 2) else "-"
    }),
    Max = sapply(params$processed_data, function(x) {
      if (is.numeric(x)) round(max(x, na.rm = TRUE), 2) else "-"
    }),
    Mean = sapply(params$processed_data, function(x) {
      if (is.numeric(x)) sprintf("%.3f", mean(x, na.rm = TRUE)) else "-"
    }),
    SD = sapply(params$processed_data, function(x) {
      if (is.numeric(x)) sprintf("%.3f", sd(x, na.rm = TRUE)) else "-"
    }),
    Unique_Values = sapply(params$processed_data, function(x) {
      if (!is.numeric(x)) length(unique(na.omit(x))) else "-"
    }),
    stringsAsFactors = FALSE
  )

  # Replace NA/NaN with dash
  summary_df[] <- lapply(summary_df, function(x) {
    ifelse(is.na(x) | x == "NaN", "-", as.character(x))
  })

  kable(summary_df, caption = "Data Summary", align = "c") %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = TRUE
    ) %>%
    scroll_box(width = "100%", height = "400px")
} else {
  cat("No data available for summary statistics.")
}
```

## User-Generated Visualizations

```{r fig.width=10, fig.height=6}
# Generate visualizations from history
if (params$include_plots && !is.null(params$preprocessing_history) &&
  !is.null(params$processed_data)) {
  # Handle list structure
  history_list <- if (is.data.frame(params$preprocessing_history)) {
    lapply(1:nrow(params$preprocessing_history), function(i) as.list(params$preprocessing_history[i, ]))
  } else {
    params$preprocessing_history
  }

  # Filter visualization operations
  viz_count <- 0
  processed_viz_signatures <- c()
  for (i in seq_along(history_list)) {
    op <- history_list[[i]]
    action <- if (!is.null(op$action)) op$action else if (!is.null(op[["action"]])) op[["action"]] else NA

    if (!is.na(action) && action == "Visualization") {
      viz_count <- viz_count + 1

      chart_type <- if (!is.null(op$method)) op$method else NA
      params_list <- if (!is.null(op$params)) op$params else if (!is.null(op[["params"]])) op[["params"]] else list()
      x_var <- if (!is.null(params_list$x)) params_list$x else NA
      y_var <- if (!is.null(params_list$y)) params_list$y else NA

      # Create unique signature for deduplication
      # Use a safe string representation for NA values in the signature
      sig_x <- if (is.na(x_var)) "NA" else x_var
      sig_y <- if (is.na(y_var)) "NA" else y_var
      sig_type <- if (is.na(chart_type)) "NA" else chart_type

      viz_sig <- paste(sig_type, sig_x, sig_y, sep = "|")

      if (viz_sig %in% processed_viz_signatures) {
        next
      }
      processed_viz_signatures <- c(processed_viz_signatures, viz_sig)

      if (!is.na(x_var) && x_var %in% names(params$processed_data)) {
        tryCatch(
          {
            # Bar chart (Simple, Top N, Ordered)
            if (!is.na(chart_type) && chart_type %in% c("bar", "topn_bar", "ordered_bar")) {
              p <- ggplot(params$processed_data, aes_string(x = x_var)) +
                geom_bar(fill = webapp_colors[1]) +
                theme_minimal() +
                theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
                labs(title = paste("Bar Chart:", x_var), x = x_var, y = "Count")
              print(p)
            }

            # Stacked bar 100%
            else if (!is.na(chart_type) && chart_type == "stacked_bar_100" && !is.na(y_var) &&
              y_var %in% names(params$processed_data)) {
              ct <- as.data.frame(prop.table(table(
                params$processed_data[[x_var]],
                params$processed_data[[y_var]]
              ), 1) * 100)
              names(ct) <- c("x_cat", "y_cat", "percent")

              p <- ggplot(ct, aes(x = x_cat, y = percent, fill = y_cat)) +
                geom_bar(stat = "identity") +
                scale_fill_manual(values = webapp_colors) +
                theme_minimal() +
                labs(
                  title = paste("Stacked Bar 100%:", x_var, "vs", y_var),
                  x = x_var, y = "Percentage (%)", fill = y_var
                )
              print(p)
            }

            # Grouped Bar
            else if (!is.na(chart_type) && chart_type == "grouped_bar" && !is.na(y_var) &&
              y_var %in% names(params$processed_data)) {
              ct <- as.data.frame(prop.table(table(
                params$processed_data[[x_var]],
                params$processed_data[[y_var]]
              ), 1) * 100)
              names(ct) <- c("x_cat", "y_cat", "percent")

              p <- ggplot(ct, aes(x = x_cat, y = percent, fill = y_cat)) +
                geom_bar(stat = "identity", position = "dodge") +
                scale_fill_manual(values = webapp_colors) +
                theme_minimal() +
                labs(
                  title = paste("Grouped Bar Chart:", x_var, "vs", y_var),
                  x = x_var, y = "Percentage (%)", fill = y_var
                )
              print(p)
            }

            # Histogram
            else if (!is.na(chart_type) && chart_type == "histogram") {
              p <- ggplot(params$processed_data, aes_string(x = x_var)) +
                geom_histogram(fill = webapp_colors[1], bins = 30) +
                theme_minimal() +
                labs(title = paste("Histogram:", x_var), x = x_var, y = "Count")
              print(p)
            }

            # Boxplot
            else if (!is.na(chart_type) && chart_type == "boxplot") {
              p <- if (!is.na(y_var) && y_var %in% names(params$processed_data)) {
                ggplot(params$processed_data, aes_string(x = x_var, y = y_var)) +
                  geom_boxplot(fill = webapp_colors[1]) +
                  labs(title = paste("Boxplot:", x_var, "vs", y_var))
              } else {
                ggplot(params$processed_data, aes_string(y = x_var)) +
                  geom_boxplot(fill = webapp_colors[1]) +
                  labs(title = paste("Boxplot:", x_var))
              }
              p <- p + theme_minimal()
              print(p)
            }

            # Pareto
            else if (!is.na(chart_type) && chart_type == "pareto") {
              counts <- table(params$processed_data[[x_var]])
              counts <- sort(counts, decreasing = TRUE)
              df <- data.frame(category = names(counts), count = as.numeric(counts))
              df$category <- factor(df$category, levels = df$category)

              p <- ggplot(df, aes(x = category, y = count)) +
                geom_bar(stat = "identity", fill = webapp_colors[1]) +
                theme_minimal() +
                theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
                labs(title = paste("Pareto Chart:", x_var), x = x_var, y = "Count")
              print(p)
            }

            # Cumulative Percent
            else if (!is.na(chart_type) && chart_type == "cumulative_percent") {
              counts <- table(params$processed_data[[x_var]])
              df <- data.frame(category = names(counts), count = as.numeric(counts))
              # Sort if needed, but usually cumulative implies ordinal or sorted
              percents <- cumsum(prop.table(counts) * 100)
              df$percent <- as.numeric(percents)

              p <- ggplot(df, aes(x = category, y = percent, group = 1)) +
                geom_line(color = webapp_colors[1], size = 1) +
                geom_point(color = webapp_colors[1]) +
                theme_minimal() +
                theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
                labs(title = paste("Cumulative Percent:", x_var), y = "Cumulative %")
              print(p)
            }
          },
          error = function(e) {
            cat("Could not generate chart:", e$message, "\n")
          }
        )
      }
    }
  }

  if (viz_count == 0) {
    cat("No visualizations were generated.")
  }
} else {
  cat("No visualizations available.")
}
```

## Correlation Analysis

```{r}
# Display correlation analysis with statistical tests
if (params$include_correlation && !is.null(params$preprocessing_history)) {
  # Source the correlation analysis script
  if (!is.null(params$script_dir)) {
    corr_script_path <- file.path(params$script_dir, "correlation_analysis.R")
    if (file.exists(corr_script_path)) {
      source(corr_script_path, local = TRUE)
    }
  }

  # Handle list structure
  history_list <- if (is.data.frame(params$preprocessing_history)) {
    lapply(1:nrow(params$preprocessing_history), function(i) as.list(params$preprocessing_history[i, ]))
  } else {
    params$preprocessing_history
  }

  corr_count <- 0
  processed_signatures <- c()

  for (i in seq_along(history_list)) {
    op <- history_list[[i]]
    action <- if (!is.null(op$action)) op$action else if (!is.null(op[["action"]])) op[["action"]] else NA

    if (!is.na(action) && action == "Correlation Analysis") {
      # Robustly extract inputs
      cols_raw <- op$inputs
      cols <- c()

      if (is.list(cols_raw)) {
        cols <- unlist(cols_raw)
      } else if (is.character(cols_raw)) {
        if (length(cols_raw) == 1 && grepl("^c\\(", cols_raw)) {
          try(
            {
              cols <- eval(parse(text = cols_raw))
            },
            silent = TRUE
          )
        } else {
          cols <- cols_raw
        }
      }

      # Deduplication check: Sort and clean columns to create unique signature
      if (length(cols) > 0) {
        # Normalize columns: distinct, sorted, trimmed
        cols <- sort(unique(trimws(as.character(cols))))

        sig <- paste(cols, collapse = "|")
        if (sig %in% processed_signatures) {
          next
        }
        processed_signatures <- c(processed_signatures, sig)
      }

      corr_count <- corr_count + 1

      method <- if (!is.null(op$method)) op$method else "Matrix"

      if (length(cols) >= 2 && all(cols %in% names(params$processed_data))) {
        cat("\n\n### Correlation Analysis: ", paste(cols, collapse = ", "), "\n\n")

        # 1. Calculate Full Correlation Matrix
        n_vars <- length(cols)
        corr_matrix <- matrix(NA, nrow = n_vars, ncol = n_vars)
        rownames(corr_matrix) <- cols
        colnames(corr_matrix) <- cols

        p_matrix <- matrix(NA, nrow = n_vars, ncol = n_vars) # Store p-values for significance stars

        # Loop through all pairs
        for (i in 1:n_vars) {
          for (j in 1:n_vars) {
            if (i == j) {
              corr_matrix[i, j] <- 1.0
              p_matrix[i, j] <- 0
            } else {
              col1 <- cols[i]
              col2 <- cols[j]

              var1 <- params$processed_data[[col1]]
              var2 <- params$processed_data[[col2]]

              # Determine types
              type1 <- if (is.ordered(var1)) "ordinal" else "nominal"
              type2 <- if (is.ordered(var2)) "ordinal" else "nominal"
              type_pair <- paste(sort(c(type1, type2)), collapse = "-")

              # Resolve method
              target_method <- "cramers_v" # Default
              op_params <- if (!is.null(op$params)) op$params else list()
              if (!is.null(op_params$methods) && !is.null(op_params$methods[[type_pair]])) {
                target_method <- op_params$methods[[type_pair]]
              } else {
                # Fallback defaults
                if (type_pair == "ordinal-ordinal") {
                  target_method <- "spearman"
                } else if (type_pair == "nominal-ordinal") target_method <- "kruskal_wallis"
              }

              # Execute Test
              stat_val <- 0
              p_val <- 1

              valid_idx <- !is.na(var1) & !is.na(var2) & var1 != "" & var2 != ""
              d1 <- var1[valid_idx]
              d2 <- var2[valid_idx]

              if (length(d1) > 0) {
                tryCatch(
                  {
                    if (exists(target_method) && is.function(get(target_method))) {
                      # Handle arg order for directional tests
                      res <- if (target_method %in% c("kruskal_wallis", "anova_eta") && type1 == "nominal" && type2 == "ordinal") {
                        do.call(target_method, list(d1, d2))
                      } else if (target_method %in% c("kruskal_wallis", "anova_eta")) {
                        do.call(target_method, list(d2, d1))
                      } else {
                        do.call(target_method, list(d1, d2))
                      }

                      if (!is.null(res) && !is.null(res$result)) {
                        # Use effect size as correlation value
                        stat_val <- if (!is.null(res$result$effect_size)) res$result$effect_size else res$result$statistic
                        p_val <- res$result$p_value
                      }
                    }
                  },
                  error = function(e) {}
                )
              }

              corr_matrix[i, j] <- stat_val
              p_matrix[i, j] <- p_val
            }
          }
        }

        # 2. Visualize Matrix (Matching Frontend Colors)

        # Convert to long format for ggplot
        corr_data <- as.data.frame(as.table(corr_matrix))
        names(corr_data) <- c("Var1", "Var2", "Correlation")

        # Add p-values for significance stars
        p_data <- as.data.frame(as.table(p_matrix))
        corr_data$PValue <- p_data$Freq

        # Add significance stars
        corr_data$Stars <- ifelse(corr_data$PValue < 0.001, "***",
          ifelse(corr_data$PValue < 0.01, "**",
            ifelse(corr_data$PValue < 0.05, "*", "")
          )
        )
        corr_data$Stars[corr_data$Var1 == corr_data$Var2] <- "" # No stars on diagonal

        # Define Color Categories matching Frontend
        # >= 0.7: Red (#EF4444)
        # >= 0.5: Orange (#FB923C)
        # >= 0.3: Yellow (#FDE047)
        # >= 0.1: Green (#BBF7D0)
        # < 0.1: Blue (#DBEAFE)

        corr_data$ColorGroup <- cut(abs(corr_data$Correlation),
          breaks = c(-1, 0.1, 0.3, 0.5, 0.7, 2),
          labels = c("Negligible", "Very Weak", "Weak", "Moderate", "Strong")
        )

        # Define colors manually
        frontend_colors <- c(
          "Negligible" = "#DBEAFE", # Blue-100
          "Very Weak" = "#BBF7D0", # Green-200
          "Weak" = "#FDE047", # Yellow-300
          "Moderate" = "#FB923C", # Orange-400
          "Strong" = "#EF4444" # Red-500
        )

        # Diagonal is gray
        corr_data$FillColor <- frontend_colors[corr_data$ColorGroup]
        corr_data$FillColor[corr_data$Var1 == corr_data$Var2] <- "#E5E7EB" # Gray-200

        # Create Heatmap
        p <- ggplot(corr_data, aes(x = Var1, y = Var2)) +
          geom_tile(aes(fill = ColorGroup), color = "white") +
          geom_text(aes(label = sprintf("%.3f", Correlation)), fontface = "bold", size = 3.5) +
          geom_text(aes(label = Stars), vjust = 2, size = 3) +
          scale_fill_manual(values = frontend_colors, name = "Correlation Strength") +
          scale_y_discrete(limits = rev(levels(corr_data$Var2))) + # Reverse Y axis to match matrix order
          labs(
            x = NULL, y = NULL, title = "Correlation Matrix Heatmap",
            caption = "* p<0.05, ** p<0.01, *** p<0.001"
          ) +
          theme_minimal() +
          theme(
            axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
            axis.text.y = element_text(face = "bold"),
            panel.grid = element_blank()
          )

        print(p)
        cat("\n\n")

        # Legend Description
        cat("**Legend:**\n")
        cat("* **Strong (≥ 0.7):** Red\n")
        cat("* **Moderate (0.5 - 0.7):** Orange\n")
        cat("* **Weak (0.3 - 0.5):** Yellow\n")
        cat("* **Very Weak (0.1 - 0.3):** Green\n")
        cat("* **Negligible (< 0.1):** Blue\n\n")
      }
    }
  }

  if (corr_count == 0) {
    cat("No correlation analysis was performed.")
  }
} else {
  cat("No correlation analysis available.")
}
```

## Executive Summary

```{r}
if (!is.null(params$processed_data) && nrow(params$processed_data) > 0) {
  cat("### Key Dataset Metrics\n\n")

  # Calculate metrics
  num_rows <- nrow(params$processed_data)
  num_cols <- ncol(params$processed_data)
  data_size <- format(object.size(params$processed_data), units = "auto")

  missing_cells <- sum(sapply(params$processed_data, function(x) sum(is.na(x))))
  total_cells <- num_rows * num_cols
  missing_pct <- round((missing_cells / total_cells) * 100, 2)

  num_numeric <- length(params$numeric_vars)
  num_categorical <- length(params$categorical_vars)

  summary_table <- data.frame(
    Metric = c(
      "Rows", "Columns", "Data Size", "Missing Data",
      "Numeric Variables", "Categorical Variables"
    ),
    Value = c(
      num_rows, num_cols, data_size, paste0(missing_pct, "%"),
      num_numeric, num_categorical
    )
  )

  kable(summary_table, caption = "Dataset Overview") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
} else {
  cat("No data available for executive summary.")
}
```
